---
title: "SWAG-R-Package Vignette"
author: "Gaetan Bakalli, Samuel Orso and Cesare Miglioli"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SWAG-R Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction 

SWAG-R is a package that train a meta-learning procedure that combines screening and wrapper methods to find a set of extremely low-dimensional attribute combinations. SWAG-R works on top of the caret package and proceeds in a forward-step manner. More specifically, it builds and tests learners starting from very few attributes until it includes a maximal number of attributes by increasing the number of attributes at each step. Hence, for each fixed number of attributes, the algorithm tests various (randomly selected) learners and picks those with the best performance in terms of training error. Throughout, the algorithm uses the information coming from the best learners at the previous step to build and test learners in the following step. In the end, it outputs a set of strong low-dimensional learners.

## Installation

First you need the devtools package installed and loaded:

```{r, eval=F}

install.packages("devtools", repos = "http://cran.us.r-project.org")

library(devtools)

```

Now you are able to get the SWAG-R directly from Github with the following code:

```{r, eval=FALSE}

install_github("SMAC-Group/SWAG-R-Package")

```

## Quick Start

The purpose of this section is to give users a general sense of the package, including the components, whatthey do and some basic usage. We will briefly go over the main functions, see the basic operations and have alook at the outputs. Users may have a better idea after this section what functions are available, which oneto choose, or at least where to seek help. More details are given in later sections.

First, we load the **SWAG-R** package:

```{r, eval=FALSE}

library(SWAG-R-Package)

```

We  a set of data created beforehand for illustration. Users can either load their own data or use those saved in the workspace.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
